% \onecolumn
\section{Abstract}
\subsection{Background}%
Outcomes for hospitalized patients, 
including metrics such as length of stay (LOS) and early readmission, 
are difficult to predict. 
Hospital systems, payers, and regulators have focused on 
improving these metrics, with uncertain benefit. 
Interpretable machine learning may assist in transparently identifying risk 
of important outcomes for individual patients and entire cohorts.
\subsection{Methods}%
We collected clinical, demographic, and institutional data 
for all hospitalizations from January 2011 to May 2018 
at a tertiary academic medical center. 
Machine learning (ML) algorithms were trained to predict 
readmission, LOS, death within 48–72 hours, and demographic data. 
Model performance was evaluated with 
area under the receiver operator characteristic curve (AUC), 
Brier score loss (BSL), and other metrics. 
Interpretations were generated using multiple feature extraction algorithms.
\subsection{Results}%
The dataset included 1,485,880 hospitalizations for 708,089 unique patients. 
Readmission rates were higher for patients who were 
older, African American, divorced/separated, on Medicare, 
and with multiple co-morbidities. 
Predictive performance was as follows: 
Thirty-day readmission, (AUC 0.76/BSL 0.1); 
LOS >5 days (AUC 0.85/BSL 0.14); 
and death within 48–72 hours (AUC 0.81/BSL x.x). 
Our ML algorithm was also able to predict 
gender (AUC 0.87/BSL 0.14); 
race (AUC 0.91/BSL 0.10); 
payer class (AUC 0.92/BSL 0.10); 
and age >65-years (AUC 0.94/BSL 0.09). 
Explanatory diagrams for individuals allowed understanding of 
clinical factors that impacted each outcome. 
\subsection{Conclusions}%
Interpretable ML algorithms can achieve 
state-of-the-art predictive power for hospital readmission and extended LOS, 
while simultaneously providing locally and globally interpretable predictions 
that account for clinical, demographic, and institutional variables.
\newpage%